%% Options:

%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[twocolumn]{ceurart}

%%
%% One can fix some overfulls
\sloppy

%%
%% Minted listings support 
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
\usepackage[finalizecache]{minted}
%% \usepackage[frozencache]{minted}
\usepackage{lmodern}
\usepackage{caption} %% corrects hyperlinks to caption
%% auto break lines
\setminted{breaklines=true}

\begin{document}

\copyrightyear{2023}
\copyrightclause{
    Copyright for this paper by its authors.
    Use permitted under Creative Commons License Attribution 4.0 
    International (CC BY 4.0).
}

\conference{
    ITAT'23: Information technologies -- Applications and Theory
    September 22--26, 2023, Tatranské Matliare, Vysoké Tatry
}

\title{Prefix-free graphs and suffix array construction in sublinear space}

\author[1]{Andrej Baláž}[
    email=andrejbalaz001@gmail.com,
]
\cormark[1]

\author[1]{Alessia Petescia}[]

\address[1]{
    Department of Applied Informatics,
    Faculty of Mathematics, Physics and Informatics,
    Comenius University, Bratislava, Slovakia
}

% Footnotes
\cortext[1]{Corresponding author}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
A recent paradigm shift in bioinformatics from a single reference genome to a pangenome brought with it several graph structures.
These graph structures must implement operations, such as efficient construction from multiple genomes and read mapping.
Read mapping is a well-studied problem in sequential data, and, together with data structures such as suffix array and Burrows-Wheeler transform, allows for efficient computation.
Attempts to achieve comparatively high performance on graphs bring many complications since the common data structures on strings are not easily obtainable for graphs.
In this work, we introduce prefix-free graphs, a novel pangenomic data structure; we show how to construct them and how to use them to obtain well-known data structures from stringology in sublinear space, allowing for many efficient operations on pangenomes.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\begin{keywords}
    computational pangenomics \sep
    graph pangenome \sep
    suffix array
\end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\input{content/01_introduction.tex}
\input{content/02_pfg.tex}
\input{content/03_sa_construction.tex}
\input{content/04_results.tex}
\input{content/05_conclusion.tex}

\section{Online Resources}
The sources for the prefix-free graphs are available via
\begin{itemize}
\item \href{https://github.com/andynet/pfg}{https://github.com/andynet/pfg}.
\end{itemize}

\begin{acknowledgments}
This research was funded by a grant from the European Union’s Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No 956229 (ALPACA) and by a grant from Slovak Research Grant Agency VEGA 1/0538/22.
\end{acknowledgments}

\bibliography{sources}

\end{document}

